Physics inspired priors for neutron star mergers improves equation of state constraints

### 2025-02-19 paper.tex
first crack at abstract and modified title

	\title{Microphysics and astrophysics inspired priors for neutron star mergers}

@@

	\begin{abstract}

	\end{abstract}

>>>>

	% \title{Microphysics and astrophysics inspired priors for neutron star mergers}
	\title{Physics inspired priors for neutron star mergers improves equation of state constraints}

@@

	\begin{abstract}
		Gravitational-wave astronomy shows great promise in determining nuclear physics in a regime not accessible to terrestrial experiments. 
		We introduce physics-inspired priors derived from a combination of constraints from nuclear-physics experiments, perturbative quantum chromodynamics calculations, and astrophysical constraints.
		When these priors are used in gravitational-wave astrophysical inference, we show significant improvement on nuclear equation of state constraints. 
		Applying these to the first two gravitational-wave binary neutron star mergers GW170817 and GW190425, constraints on the radius of a $1.4\,M_\odot$ neutron star improves from $R=\red{XX^{+YY}_{-ZZ}}\,{\rm km}$ to $\red{XX^{+YY}_{-ZZ}}\,{\rm km}$ (90\% confidence intervals).
		We also show these priors can be used to perform model selection between binary neutron star and neutron star-black hole mergers, although the results for GW190425 are incolnclusive with a Bayes factor of $\red{}$ in favour of the binary neutron star merger hypothesis, representing only marginal evidence.
		We advocate for these physics-inspired priors to be used as standard in the literature, and provide open-source code for reproducibility and further adaptation of the method.
	\end{abstract}
